Rclean: A Tool for Writing Cleaner, More Transparent Code

### 2020-01-23 joss/paper.Rmd
Edits introduction paragraph segue

From this perspective, tools that can lower the time and energy
required to re-factor and streamline analytical scripts, and otherwise
help to "clean" code will have a significant impact on scientific
reproducibility across all disciplines [@Visser2015]. To support this
objective, we have created ``Rclean`` which automatically reduces a
script to the parts that are specifically relevant to a research
product (e.g.  a blog, academic talk or research article). Here, we
detail the structure of the package's API, describe the general
workflow illustrated by an example use case and provide some
background on how the underlying use of data provenance enables the
package.

>>>>

The stability, reproducibility and re-use of scientific analyses in R
would be improved by refactoring. From this perspective, tools that
can lower the time and energy required to re-factor and streamline
analytical scripts, and otherwise help to "clean" code will have a
significant impact on scientific reproducibility across all
disciplines [@Visser2015]. More often then not, when someone is
writing an R script, the intent to produce a set of results, such as a
statistical analysis, figure, etc. This set of results is always a
subset of a much larger set of possible ways to explore a dataset, as
there are many statistical approaches and tests, let alone ways to
create visualizations and other representations of patterns in
data. This commonly leads to lengthy, complicated scripts from which
researchers manually subset results, but never refactor (i.e. reduce
to the desired subset). In part, this is driven a desire to record the
entire scripting/analytical process and not lose work, which is
alleviated by the use of proper version control systems (e.g. git and
subversion); however, refactoring will always be a part of the
analytical process regardless.

To support provide support for easier refactoring in R, we have
created ``Rclean``. The ``Rclean`` package provides tools to
automatically reduce a script to the parts that are specifically
relevant to a research product (e.g.  a scientific report, academic
talk, research article, etc.). Although potentially useful to all R
coders, it was designed to ease refactoring for scientists that use R
but do not have formal training in software engineering. Here, we
detail the structure of the package's API, describe the general
workflow illustrated by an example use case and provide some
background on how data provenance enables the underlying functionality
of the package. We then end with a discussion of future applications
of data provenance in the context of "code cleaning" and the potential
integration with other software engineering tools for the R community.
